# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data
```{r}
data <- read.csv(unz("activity.zip","activity.csv"),stringsAsFactors = FALSE)
head(data)
str(data)
```

## What is mean total number of steps taken per day?

```{r}
library(plyr)
library(ggplot2)
```

###  Histogram of the total number of steps taken each day
```{r graphic, fig.width=20}
summarized_table <- ddply(data,.(date),
                          summarise,total_steps = sum(steps,na.rm = TRUE))

steps_hist <- ggplot(summarized_table,aes(x=date)) + 
        geom_histogram(aes(weights = total_steps)) + 
        ylab("Total steps for each day") +
        theme(axis.text.x = element_text(colour = "black",angle = 90,size = 12),
              axis.text.y = element_text(colour = "black",size = 11),
              axis.title = element_text(size = 20),
              plot.title = element_text(size = 30))
steps_hist 

```
###  Mean and median total number of steps taken per day
```{r}
step_mean <- mean(summarized_table$total_steps,na.rm = TRUE )
step_median <- median(summarized_table$total_steps,na.rm = TRUE)
```
- The mean of total number of steps is `r step_mean`.

- The median of total number of steps is `r step_median`. 


## What is the average daily activity pattern?

``` {r graphics, fig.width=20}
inter_summar <- ddply(data, .(interval),summarise,mean = mean(steps,na.rm = TRUE))
pattern_plot <- ggplot(inter_summar,aes(interval,mean)) + 
        geom_line(size = 1,colour = "blue") +
        xlab("Interval") + 
        ylab("mean of number of steps")+
        theme(axis.text.x = element_text(colour = "black",size = 11),
              axis.text.y = element_text(colour = "black",size = 11),
              axis.title = element_text(size = 20),
              plot.title = element_text(size = 30))
pattern_plot 
```

```{r}
maxinter<- inter_summar$interval[which.max(inter_summar$mean)]

``` 
- The `r maxinter` contain the maximum number of steps.

## Imputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
num_missing <- sum(is.na(data$steps))
```
- The number of missing value is `r num_missing`.

### Devise a strategy for filling in all of the missing values in the dataset. 

- The strategy is fill the misssing data with the floored normal random number
with the mean and standard deviation each data

```{r}
## Calculate the mean and sd with each data
summarized_table2 <- ddply(data,.(date),
                          summarise,
                          mean_steps = mean(steps,na.rm = TRUE),
                          sd_steps = sd(steps,na.rm = TRUE))
rownames(summarized_table2) <- summarized_table2$date
```
### Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
inputed <- data$steps 
names(inputed) <- data$date

## Using the normal random number with mean and sd to fill the missing data 
for( i in seq_along(inputed)){
        if(!is.na(inputed[i])) next
        else{   
                if(is.na(summarized_table2[names(inputed[i]),][,2])) inputed[i] <- 0
                else inputed[i] <- floor(rnorm(1,summarized_table2[names[inputed[1]],2],
                                               summarized_table2[names[inputed[1]],3]))                
        }
}

## Combine the inputed vector to new data frame
newData <- data.frame(steps = inputed, 
                      date = data$date,
                      interval = data$interval,
                      stringsAsFactors = TRUE)

head(newData)
```

###  histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
summarized_tableN <- ddply(data,.(date),
                          summarise,total_steps = sum(steps),
                          mean_steps = mean(steps),
                          median_steps = median(steps))
head(summarized_tableN)
```
``` {r plot,fig.width=20}
steps_histN <- ggplot(summarized_tableN,aes(x=date)) + 
        geom_histogram(aes(weights = total_steps)) + 
        ylab("Total steps for each day") +
        theme(axis.text.x = element_text(colour = "black",angle = 90,size = 12),
              axis.text.y = element_text(colour = "black",size = 11),
              axis.title = element_text(size = 20),
              plot.title = element_text(size = 30))
steps_histN
```


The inputation step did not affect the summarize so much, because the mean and sd for imputing the missing data are all come from the sample mean and  sd 

## Are there differences in activity patterns between weekdays and weekends?
```{r}
week <- weekdays(as.Date(data$date),abbreviate = TRUE) %in% c("Sun","Sat")
week <- as.factor(week)
levels(week) <- c("weekdays","weekend")
data$status <- week 
inter_summarW <- ddply(data, .(interval,status),summarise,mean = mean(steps,na.rm = TRUE))
```
```{r,message=FALSE}
library(lattice)
xyplot(mean~interval|status,
       data= inter_summarW,
       type = "l",layout = c(1,2),
       ylab = "Number of steps")
```
## Guid to reproduce the result 
* put the "activity.zip" under the working directory"
* open the script by R studio 
* clik knit HTML 
* packages required: ggplot2, lattice, plyr,
* section information as follows 
```{r}
sessionInfo()
```
